AOD Collocated with MAIAC: New Variables

### 2019-08-18 R/drake_plan_nemia.Rmd
nemia drake

# CONUS Drake Plans  ------------------------------------------------------------
source(here("R/drake_funcs.R"))
nemia_plan <- drake_plan(
  # 1.  Aeroent ---------------------------------------------------------------------

@@

  # get nearest MODIS cells 
  refgrid = get_ref_grid(),
  # near MODIS cells within 300km
  refsub = get_near_cellsid(conus_aer, refgrid),
  # aeronet sites with cloest MODIS grid
  aer_nearest = get_nearest_cell(conus_aer, refgrid), # Aeronet matched to grid, 174x4
  aer_nearest_2 = remove_site_on_water(aer_nearest), # remove 2 sites on water

  # limit aeronet data by date (aer_btw)
  aer_data = get_stn_data(aod_dir = file_in(aer_files_path)),  # 26171206x56, 11G
  aer_btw = target(sel_data_bytime(aer_data, date_start, date_end),
                   transform = map(date_start = !!date_start, 
                                   date_end = !!date_end,
                                   .id = FALSE)),
  # limit aeronet data by station, could be NEMIA or CONUS 
  sel_aer_region = target(sel_data_bystation(aer_btw, aer_nearest_2),
                      transform = map(aer_btw), .id = FALSE),

  # 2. Interpolation -----------------------------------------------------------
  # so instead of running all the Aeronet data, only run what is used
  # it also join with the nearest grid id 
  wPred = target(interpolate_aod(sel_aer_region, aer_nearest_2), transform = map(sel_aer_region, .id = FALSE)),

  # 3. WPred Join MCD19 -----------------------------------------------------------
  # mcd is the large MCD19 dataset, limit to `refsub`:nearby(300km) when readin to 
  # control total lines, unable to read in all data together, too many rows
  mcd = target(read_mcd19(sat, file_in(mcd19path_CONUS), refsub), transform = map(sat = c("terra", "aqua"))),
  # join is the rolling-joined dataset with aeronet
  join = target(rolling_join(mcd_sat = mcd, aer_data_wPred = wPred),
                transform = cross(mcd, wPred, .id = mcd)),

  # select aer sites in mcd19. further limit the stations
  sel_aer = target(aer_in_mcd19(aer_nearest_2, mcd19 = join),
             transform = map(join)),
  # buffers 
  ref = target(ref_in_buffer(region_aer_m = sel_aer, refgrid),
             transform = map(sel_aer)),
  # sites
  p = target(get_site_list(region_aer_m = sel_aer), transform = map(sel_aer)),
  n = target(get_site_name(region_aer_m = sel_aer), transform = map(sel_aer)),

  # mapping 
  # this way doesn't work:
  # sub_grid = target(select_refgrid_subset(ref_list0 = ref, point0 = p, name0 = n),
  #              transform = map(ref, p, n))
  sub_grid = target(purrr_pmap(ref = ref, p = p, n = n),
               transform = map(ref, p, n, .id = FALSE)),
  # calculate new vars to obtain the dataset ready for modeling for terra and aqua, using join, mcd, sub_grid 
  new_var = target(aod_MODIS_newVars(aod_join_MODIS = join, 
                                 MODIS_all = mcd, refDT_sub = sub_grid),
               transform = map(join, mcd, sub_grid, .id = c(join))),

@@

  # 6. Report -----------------------------------------------------------
  print1 = target(report_rmse(cv_results),
                  transform = map(cv_results,.id = FALSE)),
  plot1 = target(plot_rmse(cv_results),
                 transform = map(cv_results,.id = FALSE)),
  plotSHAP1 = target(plot_SHAP(cv_results),
                 transform = map(cv_results,.id = FALSE)),
  report = rmarkdown::render(
    knitr_in("AOD_CONUS_Report.Rmd"),
    output_file = file_out("AOD_CONUS_Report.html"))
)
nemia_plan
#drake_plan_source(nemia_plan)
nemia_config <- drake_config(nemia_plan, cache = ssd_cache) # show the dependency
# vis_drake_graph(nemia_config, from = names(nemia_config$layout))
vis_drake_graph(nemia_config, ncol_legend = 0)

@@

future::plan(future::multiprocess)
options(future.globals.maxSize= 100*1024^3) 
make(nemia_plan, cache = ssd_cache,
     parallelism = "future", jobs = 20, memory_strategy = "speed")

>>>>

# Drake Plans  ------------------------------------------------------------

# load functions for drake: 
source(here("R/drake_funcs.R"))
# new plan Yang Liu
nemia_plan <- drake_plan(
  # 1.  Aeroent ---------------------------------------------------------------------

@@

  # get nearest MODIS cells 
  refgrid = get_ref_grid(),
  aer_nearest = get_nearest_cell(nemia_aer, refgrid), # Aeronet matched to grid, 174x4
  aer_nearest_NEMIA = remove_site_on_water(aer_nearest), # remove 2 sites

  # limit aer sites by date (aer_btw)
  aer_data = get_stn_data(aod_dir = file_in(aer_files_path)),  # 26171206x56, 11G
  aer_btw = target(sel_data_bytime(aer_data, date_start, date_end),
                   transform = map(date_start = !!date_start, 
                                   date_end = !!date_end,
                                   .id = FALSE)),
  # limit by site 
  nemia = target(sel_data_bystation(aer_btw, aer_nearest_NEMIA),
                      transform = map(aer_btw), .id = FALSE),

  # 2. Interpolation -----------------------------------------------------------
  # so instead of running all the Aeronet data, only run what is used
  # it also join with the nearest grid id 
  wPred = target(interpolate_aod(nemia, aer_nearest_NEMIA), transform = map(nemia)),

  # 3. WPred Join MCD19 -----------------------------------------------------------
  # mcd is the large MCD19 dataset 
  # `mcd19path` could be revised to CONUS later 
  mcd = target(read_mcd19(sat, file_in(mcd19path)), transform = map(sat = c("terra", "aqua"))),
  # join is the rolling-joined dataset with aeronet
  join = target(rolling_join(mcd_sat = mcd, aer_data_wPred = wPred),
                transform = cross(mcd, wPred, .id = mcd)),

  sel_aer = target(aer_in_mcd19(aer_nearest_NEMIA, mcd19 = join),
             transform = map(join)),
  # buffers 
  ref = target(ref_in_buffer(nemia_aer_m = sel_aer, refgrid),
             transform = map(sel_aer)),
  # sites
  p = target(get_site_list(nemia_aer_m = sel_aer), transform = map(sel_aer)),
  n = target(get_site_name(nemia_aer_m = sel_aer), transform = map(sel_aer)),

  # mapping 
  # this way doesn't work:
  # sub = target(select_refgrid_subset(ref_list0 = ref, point0 = p, name0 = n),
  #              transform = map(ref, p, n))
  sub_grid = target(purrr_pmap(ref = ref, p = p, n = n),
               transform = map(ref, p, n, .id = FALSE)),
  # calculate new vars: the dataset ready for modeling 
  # for terra and aqua, using join, mcd, sub 
  new_var = target(aod_MODIS_newVars(aod_join_MODIS = join, 
                                 MODIS_all = mcd, refDT_sub = sub_grid),
               transform = map(join, mcd, sub_grid, .id = c(join))),

@@

  # 6. MD -----------------------------------------------------------
  print1 = target(report_rmse(cv_results),
                  transform = map(cv_results,.id = FALSE)),
  plot1 = target(plot_rmse(cv_results),
                 transform = map(cv_results,.id = FALSE)),
  plotSHAP1 = target(plot_SHAP(cv_results),
                 transform = map(cv_results,.id = FALSE)),
  report = rmarkdown::render(
    knitr_in("report.Rmd"),
    output_file = file_out("report.html"))
)
nemia_plan
#drake_plan_source(nemia_plan)
nemia_config <- drake_config(nemia_plan, cache = ssd_cache) # show the dependency
# vis_drake_graph(nemia_config, from = names(nemia_config$layout))
vis_drake_graph(nemia_config, ncol_legend = 0)

@@

future::plan(future::multiprocess)
options(future.globals.maxSize= 100*1024^3) 
make(nemia_plan, cache = ssd_cache,
     parallelism = "future", jobs = 20)

### 2019-08-18 R/non_drake_code_and_md_files/04_AOD_join_MAIAC.Rmd
nemia drake

```{r}
refDT_sub <- readRDS("~/Intermediate/AOD_data/190724_refDT_Site_idinBuffer.rds")
# load the aeronet join MCD19
aod_lst_terra <- readRDS("~/Intermediate/AOD_data/190629_LSTjoinMCD19_t_a.rds")$aod_lst_terra
# the original MCD19 
lst_terra <- read.lst(sat = "terra", columns0 = c("idLSTpair0", "overpass_time", "Optical_Depth_047"))

@@

  # join back to the MODIS file 
  MODIS_all[, day:=format(join_time, "%Y-%m-%d")]
  setnames(MODIS_all, c("idLSTpair0", "Optical_Depth_047"), c("idLSTpair0_in_buffer", "AOD_470_buffer"), skip_absent = T)
  setkey(MODIS_all, idLSTpair0_in_buffer, day)
  setkey(aod_join_buffer, idLSTpair0_in_buffer, day)
  # aod_join_buffer2 get AOD for all the buffer grid cells
  aod_join_buffer2 <- MODIS_all[,.(idLSTpair0_in_buffer, day, AOD_470_buffer)][aod_join_buffer]

@@

## Terra 
aod_lst_terra <- aod.MODIS.newVars(aod_lst_terra, lst_terra)
head(aod_lst_terra)
# AOD - AOD mean, do later 
saveRDS(aod_lst_terra, "~/Intermediate/AOD_data/190724_LSTjoinMCD19_terra_newVars.rds")

## Aqua
aod_lst_aqua <- readRDS("~/Intermediate/AOD_data/190629_LSTjoinMCD19_t_a.rds")$aod_lst_aqua
aod_lst_aqua <- aod.MODIS.newVars(aod_lst_aqua, lst_aqua)
head(aod_lst_aqua)
saveRDS(aod_lst_aqua, "~/Intermediate/AOD_data/190724_LSTjoinMCD19_aqua_newVars.rds")

@@

```{r}
# some summary
library(ggplot2)
plot.scatter.diagonal(aod_lst_terra, "Mean_AOD10km", "AOD_470nm")
cor(as.matrix(aod_lst_terra[,c(grep("Mean_AOD", names(aod_lst_terra), value = T), "AOD_470nm"), with = F]))
pmissing <- grep("pNon", names(aod_lst_terra), value = T)
summary(aod_lst_terra[,..pmissing])
```

>>>>

```{r}
refDT_sub <- readRDS("~/Intermediate/AOD_data/190724_refDT_Site_idinBuffer.rds")
# load the aeronet join MCD19
aod_lst_terra <- readRDS("~/Intermediate/AOD_data/190629_LSTjoinMCD19_t_a.rds")$aod_lst_terra
# the original MCD19 
lst_terra <- read_lst(sat = "terra")

@@

  # join back to the MODIS file 
  MODIS_all[, day:=format(join_time, "%Y-%m-%d")]
  setnames(MODIS_all, c("nearest_refgrid", "Optical_Depth_047"), c("idLSTpair0_in_buffer", "AOD_470_buffer"), skip_absent = T)
  setkey(MODIS_all, idLSTpair0_in_buffer, day)
  setkey(aod_join_buffer, idLSTpair0_in_buffer, day)
  # aod_join_buffer2 get AOD for all the buffer grid cells
  aod_join_buffer2 <- MODIS_all[,.(idLSTpair0_in_buffer, day, AOD_470_buffer)][aod_join_buffer]

@@

## Terra 
aod_lst_terra <- aod.MODIS.newVars(aod_lst_terra, lst_terra)
head(aod_lst_terra)
dim(aod_lst_terra)
# AOD - AOD mean, do later 
# saveRDS(aod_lst_terra, "~/Intermediate/AOD_data/190724_LSTjoinMCD19_terra_newVars.rds")

## Aqua
aod_lst_aqua <- readRDS("~/Intermediate/AOD_data/190629_LSTjoinMCD19_t_a.rds")$aod_lst_aqua
aod_lst_aqua <- aod.MODIS.newVars(aod_lst_aqua, lst_aqua)
head(aod_lst_aqua)
# saveRDS(aod_lst_aqua, "~/Intermediate/AOD_data/190724_LSTjoinMCD19_aqua_newVars.rds")

@@

```{r}
# some summary
library(ggplot2)
library(SHAPforxgboost)
aod_lst_terra <- readRDS("~/Intermediate/AOD_data/190724_LSTjoinMCD19_terra_newVars.rds")
scatter.plot.diagonal(aod_lst_terra, x = "Mean_AOD10km", y = "AOD_470nm")
cor(as.matrix(aod_lst_terra[,c(grep("Mean_AOD", names(aod_lst_terra), value = T), "AOD_470nm"), with = F]))
pmissing <- grep("pNon", names(aod_lst_terra), value = T)
summary(aod_lst_terra[,..pmissing])
```
